Research log

### 2025-11-04 docs/paper.Rmd
fix wrong p value, knit to word

classoption       : "man"
output            : bookdown::gitbook

@@

One possible account of the reduced decision entropy in pretend games is that it reflects pretenders’ miscalibrated intuitions about randomness, conforming to a prototype of randomness that is itself too ordered. If the same prototype of randomness is used by pretenders to determine the number of unsuccessful guesses per game, the two measures should be correlated across participants. Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`). This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`). We interpret this as evidence that the reduction in variance reflects miscalibrated beliefs not only about randomness, but also about participants’ behaviour under a counterfactual knowledge state. Those players who thought they would strictly follow a particular sequence of guesses (low entropy), ended up producing games of more variable lengths, as their success depended more on luck. Other players adjusted their decision strategy more flexibly, perhaps attempting to produce games that are not too long or short, in line with their intuitions about randomness.

>>>>

classoption       : "man"
output            : papaja::apa6_word

@@

One possible account of the reduced decision entropy in pretend games is that it reflects pretenders’ miscalibrated intuitions about randomness, conforming to a prototype of randomness that is itself too ordered. If the same prototype of randomness is used by pretenders to determine the number of unsuccessful guesses per game, the two measures should be correlated across participants. Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`). This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`). We interpret this as evidence that the reduction in variance reflects miscalibrated beliefs not only about randomness, but also about participants’ behaviour under a counterfactual knowledge state. Those players who thought they would strictly follow a particular sequence of guesses (low entropy), ended up producing games of more variable lengths, as their success depended more on luck. Other players adjusted their decision strategy more flexibly, perhaps attempting to produce games that are not too long or short, in line with their intuitions about randomness.
